Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection

### 2017-03-08 ccppabc/text/revision/main.tex
more section 3.4 edits
Added a sentence to explain advantage of ABC when there are systematic uncertainties!
Some minor corrections including separating the ***extreme data situation*** paragraph into two since it was long!

{\bf \color{dred}
In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
posteriors for each of the parameters from both inference methods are comparable 
and consistent with the ``true'' HOD parameters. However, we note that there are 
minor discrepancies between the maringalized posterior distributions. In particular, 
the distribution for $\alpha$ derived from ABC-PMC is less biased than the $\alpha$ 
constraints from the Gaussian pseudo-likelihood approach. 
}
{\bf \color{darkgreen}
@MJV: Could you take a look at what's written above and see if you feel that it's sufficient? 
}
%The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.  In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure \ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.

@@

{\bf \color{dred} 
Accurate estimation of the covariance matrix 
in LSS, however, faces a number of challenges. It is both labor and computationally 
expensive and dependent on the accuracy of simulated mock catalogs, known to be 
unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
when inferring parameters from a Gaussian-distributed data set, using covariance 
matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
function that is {\em no longer} Gaussian.
ABC-PMC does not depend on a covariance matrix estimate; hence, it does not face 
these problems. In principle, forward models of the ABC-PMC method also have the 
advantage that they can account for sources of systematic uncertainties that affect 
observations.
}
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data. While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).  Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.

{\bf \color{dred}
Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
is agnostic about the functional form of the underlying distribution of the 
summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). 
}
{\bf \color{darkgreen}
CHH: Need a sentence here that reiterates that the Gaussian functional form 
assumption is incorrect and especially incorrect for the GMF observable case.
}
{\bf \color{dred}
In the case where the observable $\gmf(N)$ is used as a summary statistic
of a galaxy catalog, assuming a Gaussian functional form for the likelihood
would clearly misreprenset the true likelihood function. In fact, this 
discrepancy may explain the less biased constraints from the ABC-PMC analysis 
compared to the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
}

{\bf \color{dred} It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC. Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects. This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift. Another example could be a catalog with a signal-to-noise ratio cut.}  {\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest. Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.  
%Hierarchical models can modify the likelihood functions through by %taking into account the selection effects. 
Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc. }{
\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
%%%

>>>>

{\bf \color{dred}
In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
posteriors for each of the parameters from both inference methods are comparable 
and consistent with the ``true'' HOD parameters. However, we note that there are 
minor discrepancies between the maringalized posterior distributions. In particular, 
the distribution for $\alpha$ derived from ABC-PMC is less biased than the $\alpha$ 
constraints from the Gaussian pseudo-likelihood approach. 
}
{\bf \color{darkgreen}
@MJV: Could you take a look at what's written above and see if you feel that it's sufficient? 
}
{\bf \color{darkgreen}
@CHH: This looks sufficient!
}
%The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.  In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure \ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.

@@

{\bf \color{dred} 
Accurate estimation of the covariance matrix 
in LSS, however, faces a number of challenges. It is both labor and computationally 
expensive and dependent on the accuracy of simulated mock catalogs, known to be 
unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
when inferring parameters from a Gaussian-distributed data set, using covariance 
matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
function that is {\em no longer} Gaussian.
ABC-PMC does not depend on a covariance matrix estimate; hence, it does not face 
these problems. In principle, forward models of the ABC-PMC method also have the 
advantage that they can account for sources of systematic uncertainties that affect 
observations. 
} {\bf \color{darkgreen} MJV: Added this sentence to emphasize why ABC is advantageous when systematic are present. @CHH: Can you take a look at this?} {\bf \color{dred} The systematic uncertainties may affect the likelihood in an unknown way, but their impact can be simulated and marginalized out in the forward generative model of the ABC-PMC method.}
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data. While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).  Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.

{\bf \color{dred}
Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
is agnostic about the functional form of the underlying distribution of the 
summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). 
}
{\bf \color{darkgreen}
CHH: Need a sentence here that reiterates that the Gaussian functional form 
assumption is incorrect and especially incorrect for the GMF observable case.
}
{\bf \color{dred}
In the case where the observable $\gmf(N)$ is used as a summary statistic
of a galaxy catalog, assuming a Gaussian functional form for the likelihood
would clearly misreprenset the true likelihood function. In fact, this 
discrepancy may explain the less biased constraints on $\alpha$ from the ABC-PMC analysis 
compared to the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
}{\bf \color{darkgreen} @CHH : added $\alpha$ to the previous sentence since the other constraints our consistent!}

{\bf \color{dred} It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC. Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects. This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift. Another example could be a catalog with a signal-to-noise ratio cut.}

{\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest. Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.  
%Hierarchical models can modify the likelihood functions through by %taking into account the selection effects. 
Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc. }{
\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
%%%
